Impact of grids and dynamical cores in CESM2.2 on the surface mass balance of the Greenland Ice Sheet

### 2022-05-10 arcticgrids.tex
fix figs

The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution. Figure~\ref{fig:grisdx}b shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids in this study. The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid. The $\tt ne30pg3$ grid has an average $\Delta x=111.2$ km, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60$ km. Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS. The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km, and is about 10 times more expensive than the $1^{\circ}$ models.  The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.

@@

We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget. However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx}). CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}. The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids. The area errors in Figure~\ref{fig:grisdx} may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.

@@

Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx}). Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid. While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term. This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask. However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=145mm]{OMEGA500F2000climo.pdf}
\end{center}
\caption{One year averages of vertical pressure velocity at 500hPa ($\tt OMEGA500$) using (a) CAM-FV (Finite-Volume dynamical core) and (b-f) various versions of the spectral-element (SE) dynamical core at approximately $1^{\circ}$ horizontal resolution and using 32 levels. (b) is equivalent to the CESM2.0 version of the SE dynamical core using the "traditional"/"old" discretization of the pressure-gradient force (PGF). Plot (c) is equivalent to configuration (b) but using the Exner form of the PGF. Plot (d) is the same as configuration (c) but also subtracting reference profiles from pressure and temperature before applying hyperviscosity operators (which is equivalent to the CESM2.2 version of SE in terms of the dynamical core). Plots (e) and (f) are equivalent to (c) and (d), respectively, by using the SE-CSLAM ($\tt ne30pg3$) version of the SE dynamical core (i.e. separate quasi-uniform physics grid and CSLAM transport scheme).}\label{fig:OMEGA500}
\end{figure}
\begin{figure}[t]
\begin{center}
\includegraphics[width=145mm]{OMEGA500FHS94.pdf}
\end{center}
\caption{Same as Figure \ref{fig:OMEGA500} but using modified Held-Suarez forcing and the average is over 18 months (excl. spin-up).}\label{fig:OMEGA500FHS94}
\end{figure}

>>>>

The accuracy of the simulated surface mass balance is expected to be sensitive to grid resolution. Figure~\ref{fig:grisdx}a shows the average grid spacing over the Greenland Ice Sheet (\textit{GrIS} hereafter) in all six grids in this study. The $\tt ne30pg2$ grid has the coarsest representation with an average grid spacing ($\Delta x$) of $\Delta x=160$ km, and the $\tt Arctic-GrIS$ grid has the highest resolution with an average grid spacing of $\Delta x=14.6$ km, similar to the $11$ km grid spacing of the RACMO2.3 grid. The $\tt ne30pg3$ grid has an average $\Delta x=111.2$ km, substantially coarser than the $\tt f09$ grid, with an average $\Delta x=60$ km. Although $\tt ne30pg3$ and $\tt f09$ have similar average grid spacing over the entire globe, and comparable computational costs, the convergence of meridians on the FV grid enhances the resolution over the GrIS. The $\tt Arctic$ grid has an average grid spacing of $\Delta x=27.8$ km, and is about 10 times more expensive than the $1^{\circ}$ models.  The $\tt Arctic-GrIS$ grid is about twice as expensive as the $\tt Arctic$ grid.

@@

We seek to integrate SMB components over a GrIS ice mask and to diagnose their contributions to the GrIS mass budget. However, the ice masks vary across the grids, especially in comparison to the RACMO3.2 ice mask, whose total area is about $3\%$ less than that of the reference dataset (Figure~\ref{fig:grisdx})b. CLM's dataset creation tool generates the model ice mask by mapping a high-resolution dataset to the target grid using the Earth System Modeling Framework (ESMF) first-order conservative remapping algorithm \cite{ESMF}. The figure suggests that the mapping errors are less than $1.5\%$ across the CESM2 grids. The area errors in Figure~\ref{fig:grisdx}b may  seem small, but even $1-2\%$ area differences can lead to large differences in integrated SMB \cite{HETAL2022TC}.

@@

Note that we are mapping to low-resolution grids that have larger GrIS areas than the source grids (Figure~\ref{fig:grisdx})b. Since the components of equation~\ref{eq:SMB} are not confined to the ice mask, method~\ref{lbl:label1} reconstructs the SMB over the portion of the common ice mask that is outside the ice mask on the source grid. While this may be a an acceptable way to reconstruct the mass source terms over different ice masks, ice melt is zero outside the source ice mask, and so method~\ref{lbl:label1} will underestimate the mass sink term. This underestimation is systematic in method~\ref{lbl:label2}, where all variables are exclusive to the ice mask; mapping to a lower-resolution grid will dilute a field of non-zero values over the ice mask with a field of zeros outside the ice mask. However, patch-level values for processes exclusive to the ice mask (e.g., ice melt) will on average have larger magnitudes than the the grid-mean quantities used in method~\ref{lbl:label1}.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=145mm]{figs/OMEGA500F2000climo.pdf}
\end{center}
\caption{One year averages of vertical pressure velocity at 500hPa ($\tt OMEGA500$) using (a) CAM-FV (Finite-Volume dynamical core) and (b-f) various versions of the spectral-element (SE) dynamical core at approximately $1^{\circ}$ horizontal resolution and using 32 levels. (b) is equivalent to the CESM2.0 version of the SE dynamical core using the "traditional"/"old" discretization of the pressure-gradient force (PGF). Plot (c) is equivalent to configuration (b) but using the Exner form of the PGF. Plot (d) is the same as configuration (c) but also subtracting reference profiles from pressure and temperature before applying hyperviscosity operators (which is equivalent to the CESM2.2 version of SE in terms of the dynamical core). Plots (e) and (f) are equivalent to (c) and (d), respectively, by using the SE-CSLAM ($\tt ne30pg3$) version of the SE dynamical core (i.e. separate quasi-uniform physics grid and CSLAM transport scheme).}\label{fig:OMEGA500}
\end{figure}
\begin{figure}[t]
\begin{center}
\includegraphics[width=145mm]{figs/OMEGA500FHS94.pdf}
\end{center}
\caption{Same as Figure \ref{fig:OMEGA500} but using modified Held-Suarez forcing and the average is over 18 months (excl. spin-up).}\label{fig:OMEGA500FHS94}
\end{figure}
